### 2020-05-07 paper.tex
Summarize

We will conclude this introductory section with the clarification of important terms. Section~\ref{sec:interviews} presents the interview results along with the collected war stories and formalizes three recurrent challenges. With Section~\ref{sec:discussion}, we will have a look at current and previous research regarding debugging of reactive applications as well as specific tooling. We are going to validate presented challenges against currently available tools within Section~\ref{sec:study}, show threats to validity of our work in Section~\ref{sec:threats} and finally end with an outlook on future iterations.

@@

\subsection{ReactiveX and RxJS}
Reactive Extensions is an open source project providing an API description for RP concepts. It describes itself as ``\dots a combination of the best ideas from the Observer pattern, the Iterator pattern, and functional programming'' \cite{reactivex}.  The API specification introduces the concept of the  \mintinline{TypeScript}{Observable} at its core. An \mintinline{TypeScript}{Observable} has the ability to be composed with other \mintinline{TypeScript}{Observable}s and eventually describe the dependency graph known in RP. A multitude of operators allow the transformation of values and \mintinline{TypeScript}{Observable}s within the graph.

@@

In order to prepare to give an answer on the first research question, we started with a series of informal interviews. We talked to professionals about their experience in working with RP in their daily jobs. We did not limit the selection of our interview partners by the tech stack defined through our research area since we were interested in as many different perspective on the topic as possible in this early stage. Eventually, we interviewed five engineers: Four work or worked more recently with Angular\footnote{\url{todo}} and ngrx\footnote{\url{todo}} to develop frontend applications. Two of them stated to have more explicit experience with RxJS further. The fifth subject works currently with a combination of akka\footnote{\url{todo}} and akka-streams\footnote{\url{todo}} in Scala to create backend applications. The agenda available in Appendix~\ref{sec:interviewquestions} was used to maintain focus during the interview sessions.

@@

We could verify gathered interview data with the results from the new reports: Subjects tend to add log statements to their code in order to understand its behavior, especially when things get more complex. Three subjects even claimed using specialized logging functionalities provided through rxjs-spy\footnote{url{todo}}. This library allows to ``tag'' specific points in the \mintinline{TypeScript}{Observable} graph with a probe. In addition to report values traversing the graph, each probe notifies about events when an observer subscribes or unsubscribes to them or when the source \mintinline{TypeScript}{Observable} completes \cite{rxjs-lifecycle} \cite{rxjs-spy-events}. As the subjects stated, the knowledge about these additional events helps in situations where they need to find problems within a graph composing multiple asynchronous data sources on multiple levels. Without it, they were blind to what actually happens within the system.

@@

\subsection{Main Challenges}

\begin{itemize}
	\item We answer RQ1 here
	\item Present results of own informal interviews
	\item Combine with war stories
	\item Extract 3 main challenges
\end{itemize}

>>>>

We will conclude this introductory section with the clarification of important terms. Section~\ref{sec:interviews} presents the interview results along with the collected war stories and formalizes two recurrent challenges. With Section~\ref{sec:discussion}, we will have a look at current and previous research regarding debugging of reactive applications as well as specific tooling. We are going to validate presented challenges against currently available tools within Section~\ref{sec:study}, show threats to validity of our work in Section~\ref{sec:threats} and finally end with an outlook on future iterations.

@@

\subsection{ReactiveX and RxJS}

Reactive Extensions is an open source project providing an API description for RP concepts. It describes itself as ``\dots a combination of the best ideas from the Observer pattern, the Iterator pattern, and functional programming'' \cite{reactivex}.  The API specification introduces the concept of the  \mintinline{TypeScript}{Observable} at its core. An \mintinline{TypeScript}{Observable} has the ability to be composed with other \mintinline{TypeScript}{Observable}s and eventually describe the dependency graph known in RP. A multitude of operators allow the transformation of values and \mintinline{TypeScript}{Observable}s within the graph.

@@

In order to prepare to give an answer on the first research question, we started with a series of informal interviews. We talked to professionals about their experiestack defined through our research area since we were interested in as many different perspective on the topic as possible in this early stage. Eventually, we interviewed five engince in working with RP in their daily jobs. We did not limit the selection of our interview partners by the tech neers: Four work or worked more recently with Angular\footnote{\url{todo}} and ngrx\footnote{\url{todo}} to develop frontend applications. Two of them stated to have more explicit experience with RxJS further. The fifth subject works currently with a combination of akka\footnote{\url{todo}} and akka-streams\footnote{\url{todo}} in Scala to create backend applications. The agenda available in Appendix~\ref{sec:interviewquestions} was used to maintain focus during the interview sessions.

@@

We could verify previously gathered interview data with the results from the new reports: The subjects tend to manually add log statements to their code in order to understand its behavior, especially when things get more complex. Three subjects even claimed using specialized logging functionalities provided through rxjs-spy\footnote{url{todo}}. This library allows to ``tag'' specific points in the \mintinline{TypeScript}{Observable} graph with a probe. In addition to report values traversing the graph, each probe notifies about events when an observer subscribes or unsubscribes to them or when the source \mintinline{TypeScript}{Observable} completes \cite{rxjs-lifecycle} \cite{rxjs-spy-events}. As the subjects stated, the knowledge about these additional events helps in situations where they need to find problems within a graph composing multiple asynchronous data sources on multiple levels. Without it, they were blind to what actually happens within the system.

@@

\subsection{Summary}

The interpretation of the combined interview and war stories data gives us an idea about two main challenges software engineers have to overcome during debugging activities of RxJS-based code:

It is crucial to the engineers to \textbf{understand data flows}  within the \mintinline{TypeScript}{Observable} graph. Questions like ``Which source emitted a value at what time?'' or ``What is the result of a specific transformation?'' are common in this context. Once a program fails, software engineers need to \textbf{rely on traditional debugger tools}. Though providing help to some extent, the tools lack of specific knowledge about the RP runtime environment adds some oddities to the debugging process: The (i) functionalities allowing step-wise navigation through the code does not respect the RP context. Assume we have set a breakpoint at \mintinline{TypeScript}{i * 2} on line~2 of Listing~\ref{lst:rxjs-step-controls}. Once the debugger hits the breakpoint and we instruct it to ``Step Over'', the debugger will lead us to the internal implementation of the \mintinline{TypeScript}{map} operator instead to line~3. (ii) Stack traces include information about the invocation of internal functions which, most of the time, do not provide any valuable information to the engineers further.

\begin{listing}[H]
	\begin{minted}{TypeScript}
		timer().pipe(
		  map(i => i * 2),
		  map(i => i - 1)
		).subscribe();
	\end{minted}
	\caption{Example RxJS Stream With Breakpoint on Line 2 Within the Arrow Function}
	\label{lst:rxjs-step-controls}
\end{listing}
